
\begin{thebibliography}{9}
    \bibitem{RePBubLik+&ShuffLik}
    Shahrzad Haddadan, Cristina Menghini, Matteo Riondato, Eli Upfal (2022) \emph{Reducing polarization and increasing diverse navigability in graphs by inserting edges and swapping edge weights. In Proceedings of ACM WSDM'21.}
    \href{https://link.springer.com/article/10.1007/s10618-022-00875-8}{\url{https://link.springer.com/article/10.1007/s10618-022-00875-8}}
    
    \bibitem{Github}
    Cristina Menghini (2021) \emph{Reducing Polarization and Improving Diverse Navigability}
     \href{https://github.com/CriMenghini/RePBubLik}{\url{https://github.com/CriMenghini/RePBubLik}}
    
    \bibitem{Amazon}
    Amazon (2006) \emph{Amazon product co-purchasing network metadata.}
    \href{https://snap.stanford.edu/data/amazon-meta.html}{\url{https://snap.stanford.edu/data/amazon-meta.html}}
    
    \bibitem{PolBlogs}
    PolBlogs (2005) \emph{A directed network of hyperlinks between weblogs on US politics.}
    \href{http://www-personal.umich.edu/~mejn/netdata/}{\url{http://www-personal.umich.edu/~mejn/netdata/}} 
    
    \bibitem{ROV}
    Kiran Garimella, Gianmarco De Francisci Morales, Aristides Gionis, and Michael Mathioudakis (2017) \emph{Reducing Controversy by Connecting Opposing Views. In Proceedings of the Tenth ACM International Conference on Web Search and Data Mining (WSDM'17).}
    \href{https://arxiv.org/abs/1611.00172}{\url{https://arxiv.org/abs/1611.00172}}

    \bibitem{Node2Vec}
    Aditya Grover, Jure Leskovec (2016) \emph{node2vec: Scalable feature learning for networks. In Proceedings of the 22nd ACM SIGKDD international conference on Knowledge discovery and data mining.}
    \href{https://cs.stanford.edu/~jure/pubs/node2vec-kdd16.pdf}{\url{https://cs.stanford.edu/~jure/pubs/node2vec-kdd16.pdf}}
    
    \bibitem{Grafo}
    Wikipedia (2023) \emph{Graph theory}
    \href{https://en.wikipedia.org/wiki/Graph_theory}{\url{https://en.wikipedia.org/wiki/Graph_theory}}

    \bibitem{RWCC}
    Wikipedia (2022) \emph{Random walk closeness centrality}
    \href{https://en.wikipedia.org/wiki/Random_walk_closeness_centrality}{\url{https://en.wikipedia.org/wiki/Random_walk_closeness_centrality}}
\end{thebibliography}

